A Search for Supernova Remnants in NGC\,6946

### 2018-11-04 ms.tex
Frank's updates

\documentclass[modern]{aastex61}
%\documentclass{aastex5}

@@

\gal\ is a nearby \cite[6.72$\pm$ 0.15 Mpc,][]{tikhonov14}, nearly face-on \cite[32.6$\degr$,][]{deblok08} galaxy with four flocculent spiral arms.  The galaxy is currently undergoing a major starburst, and as a result it has been the site of ?? historical SNe. According to ?? the total star formation rate is ??.  A total of 121 bubbles, thought to be created by stellar winds and multiple SN at the sites of some of this SFR have been identified in H1 gas that extends well outside the bright portions of the optical galaxy \citep{boomsma08}.  The high SFR in NGC6946 is thought to be bar-driven.   Given these properties, there should be a large number of SNRs to be found in NGC6946.

Optically, SNRs are usually identified on the basis of high \sii:\ha\ ratios compared to HII regions.  In bright H II regions, most S is found in the form of S$^{++}$ or S$^{+++}$, and as a result the \sii:\ha\ ratios is 0.1 or smaller.  In SNRs, where emission is shock driven, S$^+$ is found in an extended recombination zone behind the shock and the \sii:ha ratio is typically $\geq$0.4.  The situation becomes more complicated in lower surface brightness HII regions, as recently discussed by \cite{long18} for the case of M33.

The first optical search for SNRs in NGC6946 was made by \cite{matonick97}, hereafter MF97 who used interference filter imagery to identify 27 emission nebulae with \sii:\ha\ ratios $\geq$ 0.45 as SNRs.  One of these sources, MF97-16, was, later associated with the Ultraluminous Black Hole X-ray binary NGC6946 X-1 \cite{roberts03}. Though very rare ULXs have  the hard X-ray spectra  that produce line ratios in the surrounding circumstellar ISM that resemble that expected of SNRs. To our knowledge, no other optical searches for SNRs in NGC6946 exist, nor have spectra of the remaining MF97 objects been obtained.  H II regions and SNRs also radiate at radio wavelengths.  SNRs typically have (steep) non-thermal spectra arising from synchrotron radiation whereas HII regions have (flat) thermal spectra arising from free-free and recombination radiation (check).  At radio wavelengths SNRs in nearby galaxies are generally identified as non-thermal radio sources associated with optical \ha\ emission, that latter requirement arising because many background radio sources also have non-thermal spectra and because generally the SNRs are unresolved in galaxies beyond the Local Group.  In the case of NGC6946,  \cite{lacey01} identified 35 radio-selected SNR candidates on the basis of these criteria, none of which overlap with the MF97 list of optical candidates.  Their candidates were located in the brightest parts of the spiral arms of NGC6946 and they suggested that the reason the two lists do not overlap is primarily due to selection effects.

Here we discuss a new, more sensitive optical search for SNRs in NGC6946 in which we identify a total of 147 SNR candidates using interference filters.  We also discuss spectroscopic observations of 102 of these candidates, which we use to verify  the ratios obtained from the imaging and to characterize our new optical sample.

@@

To select SNR candidates, we displayed continuum-subtracted WIYN images in all three emission lines, as well as \sii:\ha\  and \oiii:\ha\ ratio images, and a continuum image (to distinguish stars from nebulae), all in DS9.   We then visually inspected these to select SNR candidates based on a high  \sii:\ha\ ratio.  The initial inspection was carried out by Middlebury undergraduate Marc DeLaney; subsequently two of us (WPB and PFW) conducted independent inspections, and then we conferred to agree on a consensus list of 147 candidates, including the 27 from MF97.   In addition, we selected 51 candidates with relatively high \oiii:\ha\ ratios, in an attempt to identify young, ejecta-dominated SNRs such as Cas A in the Galaxy.   Though none of these O-selected candidates had ratios as extreme as for Cas A or  1E0102-72.3 in the Small Magellanic Cloud, we nevertheless selected some of these for follow-up spectroscopy.

@@

{\bf [the following is vestigial from M83 paper, but we will likely have similar tables and language here.]} Tables \ref{table-SNR} and \ref{table-O3} list all the SNR candidates for which we obtained spectra.  In both tables, column 1 gives the object name from  the B12, B14, and/or D10 lists; column 2 provides some alternative names for the objects, columns 3 and 4 give the position; column 5 the diameter (assuming a distance to M83 of 4.61 Mpc).  Most of the objects lie within the footprint of the WFC3 observations, and for the great majority of these we measured the diameters from the WFC3 images.  For candidates in the outer galaxy, plus a few low-surface-brightness ones within the WFC3 footprint, we used the Magellan images instead, adjusted for seeing.   Column 6 gives the galactocentric distance; and column 7 notes objects that were detected in the {\em Chandra} ACIS X-ray survey by \citet{long14}.  Column 8 gives the mask and slitlet number used for extracting the one-dimensional spectra.  Several objects were observed with more than one mask; listed here is the one with the highest signal-to-noise ratio.   As an indication that the sample with spectra is representative of the overall population of SNR candidates in M83, we show  histograms of number vs.\ diameter in Fig.~\ref{fig_histograms}a and of number vs.\ galactocentric distance in Fig.~\ref{fig_histograms}b. The singular exception is that we purposely ignored the crowded central starburst region since spectra in this region would have been too confused.

@@

We thank NASA and everyone for allowing us time to work on this

@@

>>>>

\documentclass[modern]{aastex62}
%\documentclass{aastex5}

@@

\gal\ is a nearby \cite[6.72$\pm$ 0.15 Mpc,][]{tikhonov14}, nearly face-on \cite[32.6$\degr$,][]{deblok08} galaxy with four flocculent spiral arms.  The galaxy is currently undergoing a major starburst, and as a result it has been the site of ten historical SNe since 1917, the most of any known galaxy. According to \citet{jarrett13} the total star formation rate is $3.2\, M_\sun\,{\rm yr}^{-1}$, a high rate that is thought to be bar-driven.  A total of 121 bubbles, probably created by stellar winds and multiple SN at the star-formation sites have been identified in \hi\ gas that extends well outside the bright portions of the optical galaxy \citep{boomsma08}.  %The high SFR in \gal\ is thought to be bar-driven.   
Given these properties, one expects that a large number of supernova remnants (SNRs) should be present in \gal.

Optically, SNRs are usually identified on the basis of high \sii:\ha\ ratios compared to H~II regions.  In bright H II regions, most S is found in the form of S$^{++}$ or S$^{+++}$, and as a result the \sii:\ha\ ratios is 0.1 or smaller.  In SNRs, where emission is shock driven, S$^+$ is found in an extended recombination zone behind the shock and the \sii:ha ratio is typically $\geq$0.4.  The situation becomes more complicated in lower surface brightness H~II regions, as recently discussed by \cite{long18} for the case of M33.

The first optical search for SNRs in \gal\ was made by \cite{matonick97}, hereafter MF97 who used interference filter imagery to identify 27 emission nebulae with \sii:\ha\ ratios $\geq$ 0.45 as SNRs.  One of these sources, MF97-16, was, later associated with the Ultraluminous Black Hole X-ray binary NGC6946 X-1 \cite{roberts03}. Though very rare ULXs have  the hard X-ray spectra that produce line ratios in the surrounding circumstellar ISM that resemble that expected of SNRs. To our knowledge, no other optical searches for SNRs in NGC~6946 exist, nor have spectra of the remaining MF97 objects been obtained.

H II regions and SNRs also radiate at radio wavelengths.  SNRs typically have (steep) non-thermal spectra arising from synchrotron radiation whereas H~II regions have (flat) thermal spectra arising from free-free and recombination radiation (check).  At radio wavelengths SNRs in nearby galaxies are generally identified as non-thermal radio sources associated with optical \ha\ emission, that latter requirement arising because many background radio sources also have non-thermal spectra and because generally the SNRs are unresolved in galaxies beyond the Local Group.  In the case of NGC~6946,  \cite{lacey01} identified 35 radio-selected SNR candidates on the basis of these criteria, none of which overlap with the MF97 list of optical candidates.  Their candidates were located in the brightest parts of the spiral arms of NGC~6946 and they suggested that the reason the two lists do not overlap is primarily due to selection effects.

Here we discuss a new, more sensitive optical search for SNRs in NGC~6946 in which we identify a total of 147 SNR candidates using interference filter imagery.  We also discuss spectroscopic observations of 102 of these candidates, which we use to verify the ratios obtained from the imaging and to characterize our new optical sample.

@@

To select SNR candidates, we displayed continuum-subtracted WIYN images in all three emission lines, as well as \sii:\ha\  and \oiii:\ha\ ratio images, and a continuum image (to distinguish stars from nebulae), all in DS9.   We then visually inspected these to select SNR candidates based on a high  \sii:\ha\ ratio.  The initial inspection was carried out by Middlebury undergraduate Marc DeLaney; subsequently two of us (WPB and PFW) conducted independent inspections, and then we conferred to agree on a consensus list of 147 candidates, including the 27 from MF97.   In addition, we selected 51 candidates with relatively high \oiii:\ha\ ratios, in an attempt to identify young, ejecta-dominated SNRs such as Cas A in the Galaxy.   Though none of these O-selected candidates had ratios as extreme as for Cas A or  1E0102-72.3 in the Small Magellanic Cloud, we nevertheless selected some of these for follow-up spectroscopy.  Finally, we inspected the positions of all nine\footnote{As of the 2011 observations; a tenth SN was recorded in 2017.} historical SNe in NGC\,6946 for evidence of nebular emission.

\todo[inline]{Note which of these have emission.}

The SNR candidates are listed in Table \ref{table_candidates}.

\todo[inline]{Why is ref for this table not picked up?  We need to describe the table in the text here.}

@@

\todo[inline]{It would probably be good to include a figure showing a few example spectra, as in M83 paper, Fig 3.}

Table \ref{table_snr_spec}  lists all the SNR candidates for which we obtained spectra.

\todo[inline]{The table needs to be described.  Tables 3 \& 4 need to be modified with longtable, or something, so they will carry over to multiple pages.}

\section{Discussion}
Here are some figures we may wish to show, and points we may want to discuss (note: KSL has done some of these figures on Evernote; those all show \oiii-selected objects, which we may choose to omit):
\begin{enumerate}
    \item Correlation between flux from spectra vs imaging (M83 paper, fig 5)
    \item \sii:\ha\ ratio vs \ha\ flux; \sii:\ha\ ratio vs GC distance (fig for this already here)
    \item We don't give any diameter values for objects.  WIYN data are not great for this, but may be worth something.  Also, aren't there WFC3 Pa$\beta$(?) images; are these useful for getting diameters? There are also a handful of \ha\ images (from ACS or WFC2) which should be useful.  If we are to use these, we need to add another short subsection under Observations.
    \item \hb:\ha\ ratio vs GC Distance: done in Evernote; this indicates that reddening is severe throughout \gal, and is most severe near the center.  Happily, this has little effect on the relative strengths of the most interesting lines: \oi, \ha, \nii, and \sii.
    \item Gradients: \nii:\ha\ and \sii:\ha\ vs GCD (M83 paper, fig 8)
    \item Comparison to shock models (see existing figure).  This appears to indicate abundances between that of LMC and that of the Galaxy (?)
    \item Also compare these same ratios with other galaxies? (M83 paper, Fig 10)
    \item Separate discussion of historical SNe.  SN1980K and SN2004et are both detected, and both show broad lines.  We see NO evidence for nebular  emission from the other historical SN sites.  See Fig.~\ref{historical_SNe}.
\end{enumerate}

@@

We thank NASA and everyone for allowing us time to work on this.  PFW acknowledges additional support from the NSF through grant AST-1714281.

@@

\begin{figure}
\plotone{spectra_trio.png}
\caption{GMOS spectra of two historical SNe we have recovered in \gal, plus one from another now-confirmed SNR candidate.  The two historical SNe show the broad emission lines characteristic of ejecta-dominated SNRs.  (this is simply a figure PFW has available) \label{historical_SNe}}
\end{figure}
